### 2014-10-16 paper/include/dtail.tex
durrHurr

Z. Cui et al. \cite{dtail} proposes DTail. They evaluates two versions of DTail referenced as DTail-R and DTail-V utilizing \textbf{R} and \textbf{V} data respectively. Even so, their main idea is that more types of refresh information can be incorporated and stored together with the \textbf{R} and \textbf{V} information. DTail can then choose the most efficient refresh schema depending on the current situation, i.e. workload and type of application. Refresh data is kept per row and is stored in the DRAM itself instead of storing it in the memory controller which many other solutions do \cite{raidr}\cite{smartrefresh}\cite{refrint}. To counter the added latency of a slower memory they utilizes spatial locality of the refresh data and performs prefetching into a FIFO buffer. As the refresh data is kept on a row level DTail can save more refreshes than other techniques that have a lower granularity.

>>>>

Cui et al. \cite{dtail} propose DTail. They evaluate two versions of DTail referenced as DTail-R and DTail-V utilizing \textit{retention time} and \textit{validity} information respectively. Even so, their main idea is that more types of refresh information can be incorporated and stored together with the \textit{retention time} and \textit{validity} information. DTail can then choose the most efficient refresh schema depending on the current situation, i.e. workload and type of application. Refresh data is kept per row and is stored in the DRAM itself instead of storing it in the memory controller which many other solutions do \cite{raidr}\cite{smartrefresh}\cite{refrint}. To counter the added latency of a slower memory they utilizes spatial locality of the refresh data and performs prefetching into a FIFO buffer. As the refresh data is kept on a row level DTail can save more refreshes than other techniques that have a lower granularity.

### 2014-10-16 paper/include/raidr.tex
durrHurr

RAIDR by J. Liu et al. \cite{raidr} divides all memory lines into different groups, called bins, based on the cell with the shortest retention time in the line. In the basic configuration there are two bins, one with a refresh rate of \(64ms - 128ms\) and one with \(128ms - 256ms\). Lines in the first bin gets refreshed every \(64ms\), lines in the second bin every \(128ms\), and lines not in any bin gets refreshed at the default refresh rate which is set to \(256ms\). The implementation is done inside the memory controller and to minimize the storage needed for keeping track of bin members they use Bloom filters. Thanks to this only 1.25 KB storage is needed for a 32 GB DRAM with the default configuration. The storage requirement can however vary depending the amount of bins used. RAIDR also accommodate for temperature variations by scaling all bins refresh rates using a timer. With RAIDR a performance improvement of 4.1\% and a decrease in power consumption by 8.3\% is achieved in a system with 32GB DRAM compared to auto-refresh.

>>>>

In this approach, RAIDR by Liu et al. \cite{raidr}, all memory rows gets divided into different groups, called bins, based on the cell with the shortest retention time in the row. In the basic configuration there are two bins, one with rows that need a refresh rate of $64ms \to 128ms$ and one with rows that need to be refreshed every $128ms \to 256ms$. Rows in the first bin gets refreshed every $64ms$, rows in the second bin every $128ms$, and all remaining rows gets refreshed at the default refresh rate which is set to $256ms$. This is done using two counters, one row counter that rolls over every $64ms$, and one counter that keeps track on how many times the row counter have rolled. With this every row will get selected as refresh candidate every $64ms$, the refresh will be performed depending on which bin the row belongs to and on the second counter's value. The actual refreshes are performed by activating the row in question essentially performing a ROR.

The implementation is done inside the memory controller meaning RAIDR can not be adopted in currently existing systems. To minimize the storage needed for keeping track of bin members they use Bloom filters. Thanks to this, only 1.25 KB storage is needed for a 32 GB DRAM with the default configuration. The storage requirement can however vary depending the amount of bins used, but on the other hand the granularity of the approach becomes higher with more bins.

RAIDR conform to temperature variations by utilizing a third counter that is used to scale all bins refresh rate with the temperature. Comparing RAIDR with the somewhat similar DTail-R Lie et al. does not achieve quite as big decrease in refresh operations as Cui et al. does. This is due to DTail-R having a higher granularity as the refresh rate is kept per row and not per row group as in RAIDR. Even so, a performance improvement of 4.1\% and a decrease in power consumption by 8.3\% is achieved in a system with 32GB DRAM compared to auto-refresh.
